### 2021-05-24 manuscript/asa_template/asa_template.Rmd

---
title: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Nicole White
  thanks: The authors gratefully acknowledge computational resources and services used in this work provided by the eResearch Office, Queensland University of Technology, Brisbane, Australia.
  affiliation: School of Public Health and Social Work, QUT

- name: Thiru Balasubramaniam, Richi Nayak
  affiliation: To add

- name: Adrian Barnett
  affiliation: School of Public Health and Social Work, QUT

keywords:
- 3 to 6 keywords
- that do not appear in the title

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: references.bib
output: rticles::asa_article
header-includes:
      - \usepackage{pdflscape}
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(knitr)
library(flextable)
library(tidytext)
library(openxlsx)

g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal')
```

# Introduction

An ideal statistical analysis will use appropriate methods to create insights from the data and inform the research questions. Unfortunately many current statistical analyses are far from ideal, with many researchers using the wrong methods, misinterpreting the results, or failing to adequately check their assumptions [@2008;@Leek2017]. Some researchers take a "mechanistic" approach to statistics, copying the few methods they know regardless of their appropriateness, and then going through the motions of the analysis [@Stark2018].

Many researchers lack adequate training in research methods, and statistics is something they do with trepidation and even ignorance [@Altman1994;@King2019]. 
However, using the wrong statistical methods can cause real harm [@Altman1994;@Brown2018] and bad statistical practices are being to used abet weak science [@Stark2018].
Statistical mistakes are a key source of waste in research and partly explain the current reproducibility crisis in science [@Allison2016]. Even when the correct methods are used, many researchers fail to describe them adequately, making it difficult to reproduce the results [@Ernst2017;@Zhou2018].
Poor statistical methods might not be caught by reviewers, as they may not be qualified to judge the statistics.
A recent survey of editors found that only 23% of health and medical journals used expert statistical review for all articles [@Hardwicke2020], which was little different from a survey from 22 years ago [@Goodman1998].

There is guidance for researchers on how to write up their statistical methods and results.
The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019]. 
More detailed guideance is given by the SAMPL and EQUATOR guidelines [@Lang2013;@Altman2016] with the latter covering all apsects of the paper. Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting. 
The awareness and use of these guidelines could be improved. There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.

Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to analyse the content of statistical methods sections included in common research outputs. Results are then evaluated to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [@Wikipedia].
Use of these methods sections indicates that little thought has gone into the statistical analysis.

# Methods

\subsection{Data sources}

We used two openly available data sources to find statistical methods sections: research articles published in _PLOS ONE_ and study protocols registered on the Australian and New Zealand Clinical Trials Registry (ANZCTR). Data sources were chosen as examples of common research outputs that include descriptions of statistical methods that were used, or are planned, for analysing study outcomes.

\subsubsection{Public Library of Science (PLOS ONE)}
\label{sec:methodsPLOS}

_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail” [@PLOS]. All reviewers are asked the question: "Has the statistical analysis been performed appropriately and rigorously?", with the possible responses of "Yes", "No" and "I don't know".

Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.

All _PLOS ONE_ articles are freely accessible via the PLOS Application Programming Interface (API). This enabled us to conduct searches of full-text articles and analyse data on articles' text content and general attributes such as publication date and field(s) of research. Statistical methods sections were extracted using a two-step approach:

_Step 1_: Targeted API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted analysis-related terms, combining the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Terms could appear anywhere within the main body of the article, to account for the placement of relevant text in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume and subject classification(s).

_Step 2_: _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections. To address this, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.

<!-- Full-text XML data were combined into a single dataset organised by DOI and subsection heading(s).  -->

Data included in our analysis were downloaded on 3 July 2020. For records that did not pass Step 2, we reviewed where initial search terms appeared in the full-text, to evaluate the likely proportion of statistical methods sections that were missed.

\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
\label{sec:methodsANZCTR}

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered. All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au).

Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes. The information provided must be in English. Studies are not peer reviewed.

For the statistical methods section, researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR]. Studies are reviewed by ANZCTR staff for completeness of key information, which does not include the completeness of the statistical methods sections.

All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. 
We used all the text available in the "Statistical methods" section. We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size. These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.

Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
Some studies were first registered on the alternative trial database _clinicaltrials.gov_ and then also posted to ANZCTR. We excluded these studies because they almost all had no completed statistical methods section as this section is not included in _clinicaltrials.gov_.

Statistical methods sections were missing for some studies downloaded from ANZCTR, including a small number of sections labelled as "Not applicable", "Nil" or "None". We therefore examined if there were particular studies where this section was more likely to be missing. Analysis considered a logistic regression model estimated in the Bayesian framework [@INLA; www.r-inla.org], with missing statistical methods section (yes/no) as the dependent variable. Date, study type (observational or interventional), number of funders and target sample size were used as independent variables. Results were reported as odds ratios with 95% credible intervals (CI).

\subsection{Full-text processing}
\label{sec:methods-cleaning}

<!-- Datasets were cleaned using functions from the `textclean package' [@textclean].  -->

Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting. _R_ code used for data extraction and cleaning is available from https://github.com/agbarnett/stats_section.

Mathematical notation was converted from Unicode characters to plain text. For example, the Unicode character corresponding to $\theta$ (<U+03B8>) was replaced with ‘theta’. Common symbols outside of Unicode blocks including ‘%’ (percent) and ‘<’ (‘less-than’) were similarly converted into plain text. General formatting was removed, including carriage returns, punctuation marks, in-text references (e.g. "[42]") centred equations, and other non-ASCII characters. Text contained inside brackets was retained to maximise content for analysis, with brackets removed. Common stop words including pronouns, contractions and selected prepositions were removed. We retained selected stop words that, if excluded, may have changed the context of statistical methods being described, for example ‘between’ and ‘against’.

We compiled an extensive list of statistical terms to standardise reported descriptions of statistical methods. An initial list was compiled by calculating individual word frequencies and identifying relevant terms that appeared at least 100 times. Further terms were sourced from index searches of three statistics textbooks [@Dobson2018,@Diggle2013,@Bland2015]. Plurals (e.g., ‘chi-squares’) unhyphenated (e.g., ‘chi square’) and combined (e.g. ‘chisquare’) terms were transformed to singular, hyphenated form (e.g., ‘chi-square’). Common statistical tests were also hyphenated (e.g., ‘hosmer lemeshow’ to ‘hosmer-lemeshow’). The final list is provided in Supplementary File 1.

\subsection{Clustering algorithm}

Text from statistical methods sections was analysed using Non-Negative Matrix Factorization (NMF). NMF is an established approach for topic modelling, and provides an effective solution for text-based clustering when dealing with high-dimensional data [@kim2014algorithms,@luong2019clustering]. In this section, we outline key details of NMF and its strengths compared with other text-based clustering methods.

For $N$ studies, let $P \in R^{M \times N}$ denote a content matrix, representing text from statistical methods sections as $M$ unique terms. Text clustering algorithms for identifying common topics across studies requires $P$ to be represented with a vector space model. In our case, unique terms in $P$ are modelled using the tf\*idf (term frequency $\times$ inverse document frequency) weighting schema, to account for the relative importance of common and rare terms.

A common problem facing text clustering algorithms is the curse of dimensionality due to the high number of terms in the doc$\times$term matrix representation [@aggarwal2012mining, @sutanto2018fine]. Applying text-based methods based on distance, density or probability therefore face difficulties in high-dimensional settings [@park2018examining,@mohotti2018corpus,@mohotti2019concept]. Specifically, distances between near and far points becomes negligible [@aggarwal2012mining]. This behaviour directly affects the performance of distance-based clustering methods such as \textit{k}-means [@jain] in accurately identifying subgroups (topics) present in the data. Furthermore, sparseness associated with high-dimensional matrix representations does not allow for differentiation between topics based on density differences [@mohotti2018corpus, @mohotti2018efficient].

To address these limitations, NMF deals with high-dimensional data by mapping it to a lower-dimensional space. This mapping this achieved by approximating $P$ with two factor matrices: \( W \in R^{M \times g} \) and \( H \in R ^{N \times g} \) [@aggarwal2012mining], such that $P \approx WH^{T}$. The number of subgroups of common topic inferred from the data is given by $g$.

The matrix factorization process approximates the lower dimensional non-negative factor matrices $W$ and $H$ such that they can represent high dimensional $P$ with the least error. Estimation of $W$ and $H$ is achieved by optimising an objective function; for NMB, the Fronbeius norm is used, equivalent to minimising the sum of squares for all elements of $P$:

\begin{equation}
\label{eq:NMFobjectivefn}
\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
\end{equation}

Following estimation, \textit{H} contains the information regarding topic membership for all studies. In our case, topic membership $(1,\ldots,g)$ for a statistical methods section is inferred from the maximum coefficient value in the corresponding row of $H$, also known as the topic coherence score. For our two datasets, we applied NMF with $g=10$.

\subsection{Content analysis}

Clustering results were visualised by word clouds to summarise frequently occurring terms associated with topic membership. Whilst these results were useful for inferring common themes across statistical methods sections, they did not indicate evidence of boilerplate text beyond the use of common terms. Follow up analysis therefore considered similarities in text between sections assigned to the same topic.

We used the Jaccard similarity index to compare the statistical methods section with the highest topic coherence score with all other sections assigned to the same topic. Indices were calculated at the sentence level, to account for instances of boilerplate text within larger methods sections. We chose the Jaccard index as an easy to interpret measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either sentence (union). Instances of boilerplate text were defined by a Jaccard index of 0.9 or higher and a difference in word count of plus or minus three words.

# Results

## PLOS ONE

```{r}
load('../../results/total.records.plos.rda')
n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
perc.include = round(100*n.include/n.api)
min.perc.stats = total.records[[2]] %>% filter(percentage==min(percentage)) %>% select(volume,percentage)
max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)
```

```{r}
load('../../results/wordcount.plos.rda')
n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
p.extremes = wordcounts %>% summarise(morethan500 = round(100*mean(words>500)),lessthan50=round(100*mean(words<50)))
n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
```

```{r, echo=FALSE, fig.show='hold',fig.align='center', out.width="49%", fig.cap="\\label{fig:consort-diagrams}Search results for PLOS ONE (left) and ANZCTR (right)."}

knitr::include_graphics(c("figures/excluded_plosone.jpg","figures/excluded_anzctr_missing.jpg"))

```

```{r, echo=FALSE, fig.align='center', out.width="60%",fig.cap="\\label{fig:plos-n}PLOS summary"}

knitr::include_graphics("figures/plos_summary.jpg")
```

<!-- ![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/plos.summary.jpg) -->

```{r}

load('../../results/stats_section_subject.classifications.rda')
stats_section_subject = mutate(stats_section_subject,
                               top3 = ifelse(subject_level_number %in% 1:3,1,0))

to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')

n.top3 = to_plot %>% filter(n==Yes)
n.biology = to_plot %>% filter(subject_level_name=='Biology and life sciences') %>% pull(n)
n.earth = to_plot %>% filter(subject_level_name=='Earth sciences') %>% pull(n)
n.computer = to_plot %>% filter(subject_level_name=='Computer and information sciences') %>% pull(n)

```

API searches returned `r format(n.api,big.mark=',')` results (DOIs) (\autoref{fig:consort-diagrams}). After partial matching, `r format(n.include,big.mark=',')` (`r perc.include`%) statistical methods sections were identified. In the final sample, 95,518 (85%) DOIs returned an exact match against common section headings: 64,133 for 'statistical analysis', 13,380 for 'statistical analyses' and 13,627 for 'data analysis'.

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) in years 2013 and 2014. This trend aligned with the total number of papers published in _PLOS ONE_ over the same period. The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`). All DOIs included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications (Figure \ref{fig:plos-n}B).

Statistical methods sections had a median length of `r n.words[['med']]` words and inter-quartile range of `r n.words[['q1']]` to `r n.words[['q3']]` words. `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more. `r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had statistical methods sections with 50 words or less, equal to the length of this paragraph.

For studies excluded based on section headings, initial search terms appeared in Introduction (number), Results (number) only.  xxx appeared in the Methods section without a section heading and xxx had a non-standard section heading. [TODO].

```{r, echo=FALSE,fig.align='center', out.width="80%",fig.cap="\\label{fig:plos-clouds}Topic clouds for ten clusters for PLOS ONE"}

knitr::include_graphics("figures/ploswordclouds.jpg")

```

The topic clouds for ten clusters are in Figure \ref{fig:plos-clouds}. Frequently occurring words reflected the use of statistical software (topics 3 and 5), descriptive statistics (topic 6), group based hypothesis testing (topics 1 and 4) and definitions of statistical significance (topics 1 and 9). Statistical methods sections related to regression (topic 2) and meta-analysis (topic 7) were also identified. Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.

```{r}
load('../../results/plos.results.10topics.rda')

ftab.plos = matches %>%
  group_by(topic_id) %>% summarise(DOIs = n(),`Median word count` = median(words),Q1=quantile(words,.25),Q3=quantile(words,.75),.groups='drop') %>%
  mutate(perc = round(100*DOIs/sum(DOIs),1))

#flextable(ftab.plos)

```

Topics related to statistical software differentiated between Prism GraphPad (topic 3: n = `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 3') %>% pull(perc)`%) and SPSS (topic 5: n = `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 5') %>% pull(perc)`%). Within topic 3, the sentence "GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses" was found in 1,941 studies. Similar phrasing was identified for topic 5: "SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis", with similar text found in 441 studies.

<!-- \begin{landscape} -->
<!-- \begin{table}[] -->
<!-- \centering -->
<!-- \caption{Example boilerplate text from PLOS ONE dataset} -->
<!-- \label{tab:plos-example-boilerplate} -->
<!-- \begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}} -->
<!-- \hline -->
<!-- \textbf{Topic} & \textbf{Statistical methods section} \\ \hline -->
<!-- 1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\ -->
<!-- & Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\ -->
<!-- \hline -->
<!-- 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\ -->
<!--  & All statistical analysis was performed using Graphpad Prism software.\\ -->
<!--  \hline -->
<!-- 4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\ -->
<!-- & Data are expressed as the mean ± SEM. Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test. P values less than 0.05 were considered significant.\\ -->
<!-- \hline -->
<!-- 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\ -->
<!--   & The results are presented as the mean ± SEM. Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). The difference was considered statistically significant at P$<$0.05.\\ -->
<!--   \hline -->
<!-- 6 & All results are expressed as means ± standard deviation (± SD).\\ -->
<!-- & Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\ -->
<!-- \hline -->
<!-- 9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\ -->
<!--  & Data are presented as the mean ± SD. Statistical analysis was performed using the Students unpaired t-test. Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\ -->
<!--  \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->
<!-- \end{landscape} -->

```{r}
jaccard.matches = read.xlsx('../../results/plos.jaccard_matches.xlsx')
ftab = jaccard.matches %>% mutate(Topic = paste('Topic',topic)) %>%
  select(Topic,sentences,matches) %>%  arrange(Topic) %>%
  rename('Statistical methods text' = sentences,
         'Matching DOIs' = matches)
knitr::kable(ftab,format = "latex", caption='Example boilerplate text from PLOS ONE dataset')
```

Definitions of statistical significance featured strongly in topic 1 (n = `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 1') %>% pull(perc)`%) and topic 9 (n = `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 9') %>% pull(perc)`%). Topic 1 reflected applications of two-tailed and unpaired Student's t-tests, however, instances of boilerplate text emphasised the 5% level of statistical significance (\ref{tab:plos-example-boilerplate}). In contrast, Topic 9 focused on multiple thresholds for declaring statistical significance by asterisk: ``$^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$'', a practice that has been criticised [@Wasserstein2019].

Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used. One-way analysis of variance also featured strongly in topic 4 (n = `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 4') %>% pull(perc)`%), combined with common methods for performing post-hoc multiple comparisons. Based on the Jaccard index, 1 in 5 studies were matched to the sentence ``Significant differences were determined using analysis of variance (ANOVA) follwo by Tukey post-hoc tests for multiple comparisons''. Frequently occurring words in topic 6 (n = `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(DOIs)`; `r filter(ftab.plos,topic_id=='Topic 6') %>% pull(perc)`%) reflected mentions of descriptive statistics for summarising continuous variables, namely means with standard deviations or standard errors.

<!-- Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183). -->

\subsection{ANZCTR}

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure \ref{fig:consort-diagrams}. Of the 12,700 included studies, 9,523 (75%) had a statistical methods section. The median length of sections was 129 words with an inter-quartile range of 71 to 219 words.

Odds ratios and 95% credible intervals for study characteristics included in the logistic regression model are presented in Table \ref{tab:anzctr-missing-odds}. Observational studies were less likely to have a missing statistical methods section compared with interventional studies. Missing sections became less likely over time. Studies with more funders and a larger target sample size were less likely to have a missing statistical methods section.

\begin{table}[]
\centering
\caption{Logistic regression results for study characteristics associated with missing statistical methods sections in ANZCTR}
\label{tab:anzctr-missing-odds}
\begin{tabular}{lcc}
\hline
Variable & Odds ratio & 95$\%$ CI \\
\hline
Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
Date (per year) & 0.90 & (0.88, 0.91) \\
Number of funders & 0.80 & (0.74, 0.86) \\
Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
\hline
\end{tabular}
\end{table}

```{r, echo=FALSE, fig.align='center', out.width="70%", fig.cap="ANZCTR topic clouds."}

knitr::include_graphics("figures/anzctrwordclouds.jpg")

```

The clustering algorithm found groups that were purely sample size calculations (topic 2, n = 1834), pilot studies (topic 5, n = 834), safety/tolerability studies (topic 6, n = 524), intervention studies (topic 8, n = 1020) and repeated measures ANOVA (topic 10 = 852). Examples of boilerplate text are provided in Table \ref{tab:anzctr-example-boilerplate} .

Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS". There were cases where the exact same method section had been re-used in a different study. For example, in topic 7 (n = 333), 232 sections stated 'descriptive statistics' or 'descriptive statistics used' with no additional details provided. Other instances with expanded descriptions of methods included topic 3 (n = 1277) outlining descriptive analyses only, and topic 6 (n = 524), for hypothesis testing methods, statistical significance and software.

In other cases, text had been slightly modified to account for changes in primary and secondary outcomes. Examples of these text changes were found in topic 2 (n = 1834) and topic 4 (n = 909); identified instances related to sample size calculations for patient recruitment to different studies (see Supplementary File 3).

Since studies included in this dataset outlined planned analyses, we suspected that some studies did not specify statistical methods because they had yet to consult with a statistician. Targeted searches for the term "statistician" across all topics returned xx matching records, with examples including "A statistician employed by hospital was used" and ``Pilot study at this point will use a statistician professionally to determine sample size calculations as required''.

\begin{landscape}
\begin{table}[]
\centering
\caption{Example boilerplate text from ANZCTR dataset}
\label{tab:anzctr-example-boilerplate}
\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
\hline
\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test. Continuous data will be compared using the Student’s t-test or Mann-Whitney U test. Two sided p values of less than 0.05 will be considered statistically significant.\\
 &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant. The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate. \\ \hline
5 & Pilot study \\
 &  No formal sample size calculation was performed \\ \hline
7 & Descriptive statistics \\
 &  Descriptive statistics used \\ \hline
9 & Linear mixed models will be used to analyse the data. \\ \hline
10 & Repeated measures of ANOVA \\
 &  Pre-, during, post- and follow-up variables will be subjected to mixed methods and repeated measures analyses to determine significant changes over (group and) time. \\ \hline
\end{tabular}
\end{table}
\end{landscape}

<!-- # **Box x: Examples of boilerplate text from ANZCTR (moved to table)** -->
<!-- #  -->
<!-- # Topic 3 -->
<!-- #  -->
<!-- # "Shapiro Wilk test was used as normality test. Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed. Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test. Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed." -->

# Discussion

The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail. As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994]. This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.

A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level. This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [@McShane2019] and the common misinterpretations of p-values [@Goodman2008].

Despite the extensive array of statistical tests available, many authors are reporting the same few methods.

One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality [@Hardwicke2020].

A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used [@Weissgerber2018]. These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed. This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.

If authors shared their code then this would provide an alternative route for checking what statistical methods were used. This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency. However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code [@Serghiou2021].

Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects. This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018]. This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study. These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers. Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].

## Limitations

We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods might be adequate.

We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section. Reiterate outcomes of random sample checking here.

We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.

We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers. The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%). We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.

<!-- \section{Supplementary} -->

<!-- \subsection{PLOS} -->

<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE clustering metrics."} -->
<!-- knitr::include_graphics("figures/supplementary/supp_plosone.50clusters.jpg") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="PLOS ONE topic values."} -->
<!-- knitr::include_graphics("figures/supplementary/plos.topicvalues.jpg") -->
<!-- ``` -->

<!-- \subsection{ANZCTR} -->

<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR clustering metrics."} -->
<!-- knitr::include_graphics("figures/supplementary/supp_anzctr.50clusters.jpg") -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, fig.align='center', out.width="100%", fig.cap="ANZCTR topic values."} -->
<!-- knitr::include_graphics("figures/supplementary/anzctr.topicvalues.jpg") -->
<!-- ``` -->