### 2021-08-03 manuscript/asa_template/asa_template_almostdone.tex
updated results

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc citation processing

\usepackage{pdflscape}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a content analysis of published statistical methods sections}

  \author{
        Nicole M White \thanks{The authors gratefully acknowledge computational resources and services
used in this work provided by the eResearch Office, Queensland
University of Technology, Brisbane, Australia.} \\
    \small{Australian Centre for Health Services Innovation and Centre for Healthcare Transformation} \\
\small{School of Public Health and Social Work, Queensland University of Technology} \\
     and \\     Thiru Balasubramaniam, Richi Nayak \\
    \small{Centre for Data Science and School of Computer Science} \\
\small{Queensland University of Technology} \\
     and \\     Adrian G Barnett \\
    \small{Australian Centre for Health Services Innovation and Centre for Healthcare Transformation} \\
\small{School of Public Health and Social Work, Queensland University of Technology} \\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a textual analysis of published statistical methods sections}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract. 200 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

An ideal statistical analysis will use appropriate methods to draw
insights from the data and inform the research questions. Unfortunately
many current statistical analyses are far from ideal, with 
researchers often using the wrong methods, misinterpreting the results, or
failing to adequately check their assumptions \citep{2008, Leek2017}.
Some researchers take a ``mechanistic'' approach to statistics, copying
the few methods they know regardless of their appropriateness, and then
going through the motions of the analysis \citep{Stark2018}. Accepting this form of methodological illiteracy is at odds
with principles of scientific inquiry, yet continues to pervade published scientific research \citep{VanCalster2021}. 
This paradox has been exemplified during the COVID-19 pandemic, which has led to unprecendented levels of published research of largely 
poor quality due to biases in conduct and reporting \citep{Glasziou2020,Wynants2020}.

Many researchers lack adequate training in research methods, and
statistics is something they do with trepidation and even ignorance
\citep{Altman1994, King2019}. However, using the wrong statistical
methods can cause real harm \citep{Altman1994, Brown2018} and bad
statistical practices are being to used abet weak science
\citep{Stark2018}. Statistical mistakes are a key source of research waste and are contributing to 
the current reproducibility crisis in
science \citep{Allison2016}. Even when the correct methods are used,
many researchers fail to describe them adequately, making it difficult
to reproduce the results \citep{Ernst2017, Zhou2018}. Poor statistical
methods might not be caught by reviewers, as they may not be qualified
to judge the statistics. A recent survey of editors found that only 23\%
of health and medical journals used expert statistical review for all
articles \citep{Hardwicke2020}, which was little different from a survey
from 22 years ago \citep{Goodman1998}.

There is guidance for researchers on how to write up their statistical
methods and results. The International Committee of Medical Journal
Editors recommend that researchers should: ``Describe statistical
methods with enough detail to enable a knowledgeable reader with access
to the original data to judge its appropriateness for the study and to
verify the reported results'' \citep{ICMJE2019}. More detailed guideance
is given by the SAMPL and EQUATOR guidelines
\citep{Lang2013, Altman2016} with the latter covering all apsects of the
paper tailored to different study designs. Both of these guidelines were led by Doug Altman, who spoke often
and for many years about the need for better statistical reporting. The
awareness and use of these guidelines could be improved. There were 256
Google Scholar citations to the SAMPL paper (as at 15 March 2021) which
is a good citation statistic for most papers, but is low considering the
millions of papers that use statistical analysis.

A potential contributor to poor statistical reporting is the temptation for researchers to re-use
descriptions of statistical methods, in an effort to make their papers resemble those of their peers 
and increase perceived chances of publication  \citep{Diong2018}. As descriptions become more common, rebuke by 
reviewers and journal editors can be difficult, as their past use may be argued by researchers as offering precedent for the 
conduct of analysis within their discipline \citep{Altman2002}. Two statisticians on this paper 
(AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. To investigate the extent of this practice, we applied a text-based 
clustering method to analyse the content of published statistical methods sections.
Clustering results are then evaluated to estimate the extent that researchers are
using cut-and-paste or `boilerplate' statistical methods sections.
Boilerplate text is that ``which can be reused in new contexts or
applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the conduct and transparent reporting of statistical analysis.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\subsection{Data sources}

We used two openly available data sources to find statistical methods
sections: study
protocols registered on the Australian and New Zealand Clinical Trials
Registry (ANZCTR) and research articles published in \emph{PLOS ONE}. Data sources were chosen as examples of common
research outputs that include descriptions of statistical methods that
were planned or used for analysing research outcomes.

\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
\label{sec:methodsANZCTR}

The ANZCTR was established in 2005 as part of a coordinated global
effort to improve research quality and transparency in clinical trials
reporting; observational studies can also be registered. All studies
registered on ANZCTR are publicly available and can be searched via an
online portal (\url{https://www.anzctr.org.au}).

Details required for registration follow a standardised template
\citep{ANZCTR}, which covers participant eligibility, the
intervention(s) being evaluated, study design and outcomes. The
information provided must be in English. Studies are not peer reviewed.

For the statistical methods section, researchers are asked to provide a
brief description of all sample size calculations, statistical
methods and planned analyses, although this section is not compulsory
\citep{ANZCTR}. Studies are reviewed by ANZCTR staff for completeness of
key information, which does not include the completeness of the
statistical methods sections.

All studies available on ANZCTR were downloaded on 1 February 2020 in
XML format. For our analysis, we used all text available in the ``Statistical
methods'' section. We also collated basic information about the study
including the study type (interventional or observational), submission
date, number of funders and target sample size. These variables were
chosen as we believed they might influence the completeness of the
statistical methods section. For example, we hypothesised that larger studies and
those with funding to be more complete. We were also interested in
changes over time.

Studies prior to 2013 were excluded as the statistical methods section
appeared to be introduced in 2013. Some studies were first registered on
the alternative trial database \emph{clinicaltrials.gov} and then also
posted to ANZCTR. We excluded these studies because they almost all had
no completed statistical methods section as this section is not included
in \emph{clinicaltrials.gov}.

Statistical methods sections were missing for some studies downloaded
from ANZCTR, including sections labelled as ``Not
applicable'', ``Nil'' or ``None''. We examined if there were
particular studies where the statistical methods section was more likely to be missing.
Analysis considered a logistic regression model estimated in the
Bayesian framework {[}\citet{INLA}; www.r-inla.org{]}, with missing
statistical methods section (yes/no) as the dependent variable. The independent variables were date,
study type (observational or interventional), number of funders and
target sample size. Results were
reported as odds ratios with 95\% credible intervals (CI).

\subsubsection{Public Library of Science (PLOS ONE)}
\label{sec:methodsPLOS}

\emph{PLOS ONE} is a open access mega-journal that publishes original
research across a wide range of scientific fields. Article submissions
are handled by an academic editor who selects peer reviewers based on
their self-nominated area(s) of expertise. Currently there are 324 academic editors out of 9,648 (3\%) 
with the keywords of "statistics (mathematics)" or "statistical methods" in their expertise list (web search on 25-May-2021,
 https://journals.plos.org/plosone/static/editorial-board). Submissions do not undergo
formal statistical review. Instead, reviewers are required to assess
submissions against several publication criteria, including whether:
``Experiments, statistics, and other analyses are performed to a high
technical standard and are described in sufficient detail''
\citep{PLOS}. All reviewers are asked the question: ``Has the
statistical analysis been performed appropriately and rigorously?'',
with the possible responses of ``Yes'', ``No'' and ``I don't know''.

Authors are encouraged to follow published reporting guidelines such as
EQUATOR, to ensure that chosen statistical methods are appropriate for
the study design, and adequate details are provided to enable
independent replication of results.

Data on all \emph{PLOS ONE} articles can be accessed via the PLOS
Application Programming Interface (API). This enabled us to conduct
searches of full-text articles and analyse data on articles' text
content and general attributes such as publication date and field(s) of
research. We applied a two-step
approach to identify statistical methods sections:

\emph{Step 1}: Targeted API searches were completed using the R package
`rplos' \citep{rplos}. Search queries targeted analysis-related terms,
combining the words ``data'' or ``statistical'' with one of:
``analysis'', ``analyses'', ``method'', ``methodology'' or
``model(l)ing''. Terms could appear anywhere within the main body of the
article, to account for the placement of relevant text in different
sections, for example, in the \emph{Material and Methods} section versus
\emph{Results}. Search results were indexed by a unique Digital Object
Identifier (DOI). Attribute data collected per DOI included journal
volume and subject classification(s).

\emph{Step 2}: \emph{PLOS ONE} does not use standardised headings
to preface statistical methods sections. To address this, we performed
partial matching on available headings against frequently used terms in
initial search results: `Statistical analysis', `Statistical analyses',
`Statistical method', `Statistics', `Data analysis' and `Data analyses'. Data were downloaded on 3 July 2020.

\subsection{Full-text processing}
\label{sec:methods-cleaning}

Text cleaning aimed to standardise notation and statistical terminology,
whilst minimising changes to article style and formatting. \emph{R} code
used for data extraction and cleaning is available from
\url{https://github.com/agbarnett/stats_section}.

Mathematical notation was converted from Unicode characters to plain
text. For example, the Unicode character corresponding to \(\theta\)
(\textless U+03B8\textgreater) was replaced with `theta'. Common symbols
outside of Unicode blocks including `\%' (percent) and `\textless{}'
(`less-than') were similarly converted into plain text. General
formatting was removed, including carriage returns, punctuation marks,
in-text references (e.g.~``{[}42{]}'') centred equations, and other
non-ASCII characters. Bracketed text was retained with brackets removed to
maximise content for analysis. Common stop words
including pronouns, contractions and selected prepositions were removed.
We retained selected stop words that, if excluded, may have changed the
context of statistical methods being described, for example `between'
and `against'.

We compiled an extensive list of statistical terms to standardise
reported descriptions of statistical methods. An initial list was
compiled by calculating individual word frequencies and identifying
relevant terms that appeared at least 100 times. Further terms were
sourced from index searches of three statistics textbooks
\citep{Diggle2013,Bland2015,Dobson2018}. Plurals (e.g.,
`chi-squares') unhyphenated terms (e.g., `chi square') and combined terms
(e.g.~`chisquare') were transformed to singular, hyphenated form
(e.g., `chi-square'). Common statistical tests were also hyphenated
(e.g., `hosmer lemeshow' to `hosmer-lemeshow'). The final list is
provided in Supplementary File 1.

\subsection{Clustering algorithm}

Text from statistical methods sections was analysed using Non-Negative
Matrix Factorization (NMF). NMF is an established approach for topic
modelling, and provides an effective solution for text-based clustering
when dealing with high-dimensional data
\citep[\citet{luong2019clustering}]{kim2014algorithms}.
%In this section,
%we outline key details of NMF and its strengths compared with other
%text-based clustering methods.

For \(N\) studies, let \(P \in R^{M \times N}\) denote a content matrix of text from statistical methods sections, comprising of \(M\) unique
terms. Text clustering algorithms for identifying common topics across
studies requires \(P\) to be represented with a vector space model. In
our case, unique terms in \(P\) are modelled using the tf-idf (term
frequency \(\times\) inverse document frequency) weighting schema, to
account for the relative importance of common and rare terms.

A common problem facing text clustering algorithms is the curse of
dimensionality due to the high number of terms in the doc\(\times\)term
matrix representation
\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Applying text-based
methods based on distance, density or probability therefore face
difficulties in high-dimensional settings
\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
Specifically, distances between near and far points becomes negligible
\citep{aggarwal2012mining}. This behaviour directly affects the
performance of distance-based clustering methods such as
\textit{k}-means \citep{jain2010} in accurately identifying subgroups
(topics) present in the data. Furthermore, sparseness associated with
high-dimensional matrix representations does not allow for
differentiation between topics based on density differences
\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.

To address these limitations, NMF deals with high-dimensional data by
mapping it to a lower-dimensional space. This mapping this achieved by
approximating \(P\) with two factor matrices: \(W \in R^{M \times g}\)
and \(H \in R ^{N \times g}\) \citep{aggarwal2012mining}, such that
\(P \approx WH^{T}\). The number of subgroups of common topics inferred
from the data is given by \(g\).

The matrix factorization process approximates the lower dimensional
non-negative factor matrices \(W\) and \(H\) such that they can
represent high dimensional \(P\) with the least error. Estimation of
\(W\) and \(H\) is achieved by optimising an objective function; for
NMB, the Fronbeius norm is used, equivalent to minimising the sum of
squares for all elements of \(P\):

\begin{equation}
\label{eq:NMFobjectivefn}
\min \frac{1}{2}\|P - WH\|= \sum _{i=1}^{M}\sum _{j=1}^{N} \left(  P_{i,j} -\left(WH \right)_{i,j} \right)^{2}
\end{equation}

Following estimation, \textit{H} contains the information regarding
topic membership for all studies. In our case, topic membership
\((1,\ldots,g)\) for a statistical methods section is inferred from the
maximum coefficient value in the corresponding row of \(H\), also known
as the topic coherence score. For our two datasets, we applied NMF with
\(g=10\) topics.

\subsection{Content analysis}

Results were visualised by word clouds to summarise
frequently occurring terms associated with topic membership. Follow up analysis considered 
similarities in text between statistical methods sections assign to the same topic.

Evidence of boilerplate text was assessed at the section and sentences levels using the Jaccard similarity index. 
We chose the Jaccard index as an easy to interpret measure, which summarises the similarity between two pieces of text by the
number of words common to each text (intersection), divided by the
number of words that appeared in either text (union). For both section-level and sentence-level analyses, text was tokenised at the word level. 
Instances of boilerplate text were defined by a Jaccard index of 0.8 or higher.

To reduce the number of pairwise comparisions, locality-sensitive hashing was implementing using the R package `textreuse' \citep{textreuseR}. For all topics, locality-sensitive hashing was applied using the same random seed, 
with 1,000 randomly selected minhashes broken into 200 equal-sized bands.

A similar approach was applied at the sentence level. In this case, we aimed to identify common sentences or phrases within statistical methods sections that were representative of each topic. 
Analyses considered keyword searches using dominant terms from word clouds and the reuse of sentences within the top 1,000 sections assigned to each topic, based on estimated coherence scores.

\clearpage

\section{Results}\label{results}

\subsection{ANZCTR}

We downloaded 28,008 studies. The numbers of excluded studies are shown
in Figure \ref{fig:consort-diagrams}. Of the 12,700 included studies,
9,523 (75\%) had a statistical methods section. The median length of
sections was 129 words with an inter-quartile range of 71 to 219 words. 
Sections for TODO studies were only one word, including ``ANOVA'',
``t-test'', ``SPSS'' and even ``SSPS''.

\begin{figure}[htbp]
\centering 
\includegraphics[width=0.49\linewidth]{figures/excluded_anzctr_missing}  \includegraphics[width=0.49\linewidth]{figures/excluded_plosone}
\caption{\label{fig:consort-diagrams}Search results for ANZCTR (left) and PLOS ONE (right).}
\end{figure}

Odds ratios and 95\% credible intervals for study characteristics
associated with missing statistical methods sections are in Table
\ref{tab:anzctr-missing-odds}. Observational studies were less likely to
have a missing statistical methods section compared with interventional
studies. Missing sections became less likely over time. Studies with
more funders and a larger target sample size were less likely to have a
missing statistical methods section.

\begin{table}[]
\centering
\caption{Logistic regression results for study characteristics associated with missing statistical methods sections in ANZCTR}
\label{tab:anzctr-missing-odds}
\begin{tabular}{lcc}
\hline
Variable & Odds ratio & 95$\%$ CI \\
\hline
Study type $=$ Observational & 0.78 & (0.69, 0.89) \\
Date (per year) & 0.90 & (0.88, 0.91) \\
Number of funders & 0.80 & (0.74, 0.86) \\
Target sample size (per doubling) & 0.90 & (0.88, 0.92)\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\centering 
\includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}
\caption{Word clouds for ten topics for statistical methods sections published in ANZCTR}\label{fig:anzctr-wordclouds}
\end{figure}

Figure \ref{fig:anzctr-wordclouds} shows word clouds for the ten topics identified by the NMF algorithm. Topics reflected sample size calculations (topics 2 and 5), 
study designs (topics 4, 5, 6 and 8),
quantitative methods (topics 3, 7, 9 and 10) and qualitative methods (topic 1). 
Common statistical methods for data analysis were t-testing, descriptive statistics, linear models and repeated measures ANOVA.

\begin{table}[htbp]
\centering
\caption{Results of boilerplate analysis applied to the ANZCTR dataset. The number of studies with Jaccard scores greater than or equal to 0.8 are presented. The number of exact matches is provided in brackets.}
\label{tab:anzctr-boilerplate-matches}
\begin{tabular}{lccc}
\hline
 &  & \multicolumn{2}{c}{Boilerplate text} \\
\cline{3-4}
Topic & Total studies & Section-level & {Sentence-level}\\
\hline
1 (Qualitative methods)& 854 & 53 (24) & 288 (206)\\
2 (Sample size calculations) & 1,834 & 68 (26) & 252 (172)\\
3 (Student's t-test) & 1,277 & 105 (45) & 552 (379)\\
4 (Efficacy and safety studies) & 909 & 85 (11) & 261 (203)\\
5 (Pilot studies) & 834 &  53 (34) & 148 (114)\\
6 (Safety and tolerability studies) & 524 & 51 (29) & 241 (186)\\
7 (Descriptive analysis) & 333 & 44 (39) & 69 (60)\\
8 (Intervention studies) & 1,020 & 33 (8) & 300 (203)\\
9 (Linear models) & 1,195 & 86 (33) & 436 (333)\\
10 (ANOVA) & 852 &  64 (29) & 270 (208)\\
\hline
\end{tabular}
\end{table}

Approximately 7$\%$ of all study sections were classified as boilerplate text (n = 642). One in ten sections associated with descriptive analysis were exact matches (Table \ref{tab:anzctr-boilerplate-matches}). Evidence of boilerplate text was stronger at the sentence level, with 29$\%$ of sections having at least one sentence flagged (n = 2,817).

Boilerplate text for selected topics emphasised definitions of statistical significance, combined with planned statistical methods. In topic 3 [TODO] Other topics simply stated the main analysis method, including ``descriptive statistics/descriptive analysis'' (topic 7, n = TODO), ``linear (mixed) models'' (topic 9, n = TODO) and ``repeated measures ANOVA'' (topic 10, n = TODO). Examples of common boilerplate sentences for describing statistical methods are in Table \ref{tab:anzctr-boilerplate-examples}. Full results for all topics are provided as Supplementary material.

In other cases, text had been slightly modified to account for changes in primary and secondary outcomes for a given study design. This behaviour was reflected in the high numbers of studies with boilerplate text at the section and sentence levels (Table \ref{tab:anzctr-boilerplate-matches}). Evidence of boilerplate text was strongest for topic 6, where matches aligned with the structure:

\begin{center}
here is some \underline{text} that i changed
\end{center}

Similar examples were identified for sample size calculations:

\begin{center}
here is some \underline{text} that i changed
\end{center}

Since studies in this dataset described planned analyses, we
hypothesised that some studies did not specify statistical methods because
they had yet to consult with a statistician. Targeted searches for ``statistician'' across all topics returned TODO matching studies,
with examples including ``A statistician employed by hospital was used''
and ``Pilot study at this point will use a statistician professionally
to determine sample size calculations as required''.

%\begin{landscape}
%\begin{table}
%
%\caption{\label{tab:anzctr-example-boilerplate-sentence} Examples of boilerplate text from ANZCTR studies based on targeted keyword searches (sentence level). The number of matching papers was based on a Jaccard score of 0.75 or higher}
%\centering
%\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
%\hline
%Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
%\hline
%\multicolumn{4}{l}{Statistical methods}\\
%\hline
%3 & ``t-test'' & 131 & Continuous Normally distributed variables will be compared using Student's t-test and reported as means standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range\\
%9 & ``model'' &  339 & Linear mixed models will be used to analyse the data.\\
%10 & ``ANOVA'' & 86 & Post-hoc paired comparisons corrected for multiple comparisons will be performed if ANOVA reveal significant effects.\\
%\hline
%\multicolumn{4}{l}{Study designs}\\
%4 & ``Patient'' & 375 & At a confidence level of 95\% and a precision around the estimate of 5\% a minimum of 73 patients must be included.\\
%5 & ``Pilot study'' & 16 & This is a pilot study.\\
%6 & ``Subject' & & \\
%8 & ``'group' & & \\
%\hline
%\multicolumn{4}{l}{Study designs}\\
%2 & ``'80 percent power'
%\hline
%\multicolumn{4}{l}{Qualitative analysis}\\
%\hline
%1 & & &\\
%\hline
%\multicolumn{4}{l}{Descriptive statistics}\\
%\hline
%7 & ``Descriptive'' & 33 & Analyses will be of a descriptive nature.\\
%\hline
%
%\multicolumn{4}{l}{Experiments}\\
%\hline
%10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
%\hline
%\end{tabular}
%\end{table}
%\end{landscape}

\begin{table}
\caption{\label{tab:anzctr-example-boilerplate} Example boilerplate text from ANZCTR studies with the highest number of matches per topic (sentence level). The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words; $^{*}$ Most frequent match after excluding statements of statistical significance}
\centering
\begin{tabular}[t]{p{0.05\textwidth}p{0.8\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching papers\\
\hline
1 & Significance will be set at p $<$ 0.05 & 13\\
 &  All analyses will be conducted on an intention-to-treat basis$^{*}$ & 8\\
\hline
2 & Analysis will be by intention-to-treat & 15\\
\hline
3 & A p-value $<$ 0.05 is considered statistically significant & 39\\
 & Continuous normally distributed variables will be compared using Student's t-test and reported as means \& standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range$^{*}$ & 8\\
\hline
4 & Data will be analysed by two methods the Kaplan Meier method and per-protocol analysis & 27\\
\hline
5 & This is a pilot study only & 5\\
\hline
6 & Due to the exploratory nature of this first in human study no power or sample size calculations have been performed & 7\\
\hline
7 & Descriptive statistics & 15\\
\hline
8 & The level of statistical significance will be set at p $<$ 0.05 & 11\\
& Relationships between observations at different occasions will be modelled with an unstructured covariance matrix$^{*}$ & 9\\
\hline
9 & All analyses will be on an intention-to-treat basis & 24\\
\hline
10 & However allowing for a 50 percent attrition rate we will be required to recruit n=70 participants & 10\\

\hline
\end{tabular}
\end{table}

%\begin{landscape}
%\begin{table}[]
%\centering
%\caption{Example boilerplate text from ANZCTR dataset}
%\label{tab:anzctr-example-boilerplate}
%\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
%\hline
%Topic & Statistical methods text & Matching studies\\
%3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test. Continuous data will be compared using the Student’s t-test or Mann-Whitney U test. Two sided p values of less than 0.05 will be considered statistically significant. & \\
% &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant. The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate. & \\ \hline
%5 & Pilot study & \\
% &  No formal sample size calculation was performed & \\ \hline
%7 & Descriptive statistics & \\ \hline
%9 & Linear mixed models will be used to analyse the data. & \\ \hline
%10 & Repeated measures of ANOVA & \\ \hline
%\end{tabular}
%\end{table}
%\end{landscape}

\subsection{PLOS ONE}\label{plos-one}

API searches returned 131,847 papers (DOIs)
(\autoref{fig:consort-diagrams}). After partial matching, 111,731 (85\%)
statistical methods sections were identified. In the final sample,
95,518 (85\%) DOIs returned an exact match against common section
headings: 64,133 for `statistical analysis', 13,380 for `statistical
analyses' and 13,627 for `data analysis'.

\begin{figure}[htbp]

{\centering \includegraphics[width=0.6\linewidth]{figures/plos_summary}

}

\caption{\label{fig:plos-n} A: Search results by PLOS ONE volume; B: Subject classifications assigned to full-text recorded included in the analysis}
\end{figure}

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The
total number of API search results peaked at volumes 8 (n = 19,045) and
9 (n = 19,045) in years 2013 and 2014.  The
percentage of records that included a statistical methods section varied between 64\%
(volume 2) and 86\% (volume 9). All DOIs included ``Biology and life
sciences'' (n = 107,584), ``Earth sciences'' (n = 7,605) and/or ``Computer and
information sciences'' (n = 5,190) in their top 3 subject classifications
(Figure \ref{fig:plos-n}B).

Statistical methods sections had a median length of 127 words and
inter-quartile range of 61 to 254 words. 7,450 articles (7\%) had a
statistical methods section of 500 words or more. 19,461 articles (17\%)
had statistical methods sections with 50 words or less, equal to the
length of this paragraph.

%For studies excluded based on section headings (n = 20,116), analysis-related search terms for 2,136 DOIs appeared in the general text of the Materials and Methods section. In other cases,
 %search terms oftern appeared in the Introduction (n = 728) and Results and/or Discussion (n = 981) sections. 4,010 DOIs included at least 1 section heading that contained the terms ``analysis'' or ``analyses'' without direct reference to statistical method(s) (e.g. Microarray analysis).

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{figures/ploswordclouds}

}

\caption{\label{fig:plos-clouds}Word clouds for ten topics for statistical methods sections published in PLOS ONE}\label{fig:plos-clouds}
\end{figure}

Frequently occurring words by topic reflected the use of statistical software
(topics 3 and 5), descriptive statistics (topic 6), group based
hypothesis testing (topics 1 and 4) and definitions of statistical
significance (topics 1 and 9) (\ref{fig:plos-clouds}). Topics related to
regression (topic 2) and meta-analysis (topic 7) were also identified.

Statistical software topics differentiated between
GraphPad Prism (topic 3: n = 9,974) and SPSS (topic 5: n = 9,648).
Within topic 3, targeted searches for ``GraphPad'' identified 7,141 sentences from 6,844 papers. Of these,
1,306 papers met our definition of boilerplate text, with examples varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.

\begin{table}[htbp]
\centering
\caption{Results of boilerplate analysis applied to the PLOS ONE dataset. The number of studies with Jaccard scores greater than or equal to 0.8 are presented. The number of exact matches is provided in brackets.}
\label{tab:plosr-boilerplate-matches}
\begin{tabular}{lccc}
\hline
 &  & \multicolumn{2}{c}{Boilerplate text} \\
\cline{3-4}
Topic & Total studies & Section-level & {Sentence-level}\\
\hline
1 (Student's t-test) & 3,784 & 344 () & )\\
2 () & 19,649 & 135 () &)\\
3 (GraphPad Prism) & 9,974 & 405 () &)\\
4 (ANOVA) & 10,212 & 452 () &)\\
5 (SPSS) & 9,648 & 287 () &)\\
6 (Descriptive analysis) & 4,764 & 331 () & \\
7 (Meta analysis) & 5,278&  65 () & \\
8 () & 42,261 &  116 () & \\
9 () & 6,195 & 260 () & \\
10 () & 4,693 & 336 () & \\
\hline
\end{tabular}
\end{table}

\begin{landscape}
\begin{table}

\caption{\label{tab:plos-example-boilerplate-sentence} Examples of boilerplate text from PLOS ONE papers based on targeted keyword searches (sentence level). The number of matching papers was based on a Jaccard score of 0.75 or higher}
\centering
\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
\hline
Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
\hline
\multicolumn{4}{l}{Hypothesis testing}\\
\hline
1 & ``t-test'' & 531 & Statistical analysis was performed using Student's t-test.\\
4 & ``ANOVA'' & 270 & Statistical differences were analysed by one-way ANOVA followed by multiple comparisons performed with post-hoc Bonferroni test (SPSS version 16).\\
9 & ``p $<$'' & 2,088 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
\hline
\multicolumn{4}{l}{Statistical software}\\
\hline
3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
\hline
\multicolumn{4}{l}{Descriptive statistics}\\
\hline
6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
\hline
\multicolumn{4}{l}{Experiments}\\
\hline
10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
\hline
\end{tabular}
\end{table}
\end{landscape}

\pagebreak

%\begin{figure}
%\centering
%\includegraphics[width=0.8\linewidth]{figures/plosjaccardsectionlevel} 
%
%\caption{\label{fig:plos-clouds}Total matches by topic for statistical methods sections published in PLOS ONE (section level)}\label{fig:plos-jaccard-section}
%\end{figure}

\begin{table}

\caption{\label{tab:plos-example-boilerplate} Example boilerplate text from PLOS ONE papers with the highest number of matches per topic (sentence level). The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words; $^{*}$ Most frequent match after excluding statements of statistical significance}
\centering
\begin{tabular}[t]{p{0.05\textwidth}p{0.8\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching papers\\
\hline
1 & A p-value of less-than 0.05 was considered statistically significant & 664\\
 &  Statistical analysis was performed using student t-test$^{*}$ & 53\\
\hline
2 & A p-value less-than 0.05 was considered to be statistically significant & 1,138\\
 & All statistical analyses were performed using sas version 9.4 sas institute cary nc usa$^{*}$ & 140\\
\hline
3 & A p-value less-than 0.05 was considered significant & 914\\
 & Data are expressed as mean $\pm$ SEM$^{*}$ & 350\\
\hline
4 & P less-than 0.05 was considered statistically significant & 580\\
 &  Data are presented as the mean $\pm$ SEM$^{*}$ & 359\\
\hline
5 & A p-value less-than 0.05 was considered a significant difference & 739\\
 & All data are presented as mean $\pm$ standard deviation$^{*}$ & 149\\
\hline
6 & A p-value of less-than 0.05 was considered significant & 519\\
&  Data are presented as mean $\pm$ SEM$^{*}$ & 223\\
\hline
7 & A p-value less-than 0.05 was considered statistically significant & 305\\
 &  Otherwise the fixed-effects model was used$^{*}$ & 45\\
\hline
8 & Data were analysed using R version 2.15.2 & 10\\
\hline
9 & Significance was set at p less-than 0.05 & 195\\
& All data are presented as mean $\pm$ SEM$^{*}$ & 115\\
\hline
10 & A p-value less-than 0.05 was considered significant & 397\\
& All the experiments were repeated at least three times$^{*}$ & 119\\

\hline
\end{tabular}
\end{table}

Definitions of statistical significance featured strongly in topic 1 (n
= 3,784) and topic 9 (n = 6,195). Topic 1 reflected applications of Student's t-tests, however,
instances of boilerplate text emphasised the 5\% level of statistical significance (Table \ref{tab:plos-example-boilerplate}). Topic 9
focused on multiple thresholds for declaring statistical significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
practice that has been criticised \citep{Wasserstein2019}. Variations of this phrase were identfied in 1,676 papers.

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance featured strongly in topic 4 (n = 10,212), combined
with common methods for performing post-hoc multiple comparisons (Table \ref{tab:plos-example-boilerplate-sentence}). 
Frequently occurring words in topic 6 (n = 4764) reflected
mentions of descriptive statistics for summarising continuous variables,
namely means with standard deviations or standard errors. The sentence starting with ``data was (were) expressed as (the) mean $\pm$'' returned exact matches in 523 papers.

Statements of statistical significant were the most common form of boilerplate text across nine out of ten topics (\ref{tab:plos-example-boilerplate}). In all cases, statistical significance was defined at p$<$0.05.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The first line in many statistical analysis sections in \emph{PLOS ONE}
was the software used and some entire sections in ANZCTR only stated the
software, implying that the software is the most important detail. As
Doug Altman said, ``Many people think that all you need to do statistics
is a computer and appropriate software'' \citep{Altman1994}. This is far
from the truth, and whilst it is important for researchers to mention
the software and version used for reproducibility purposes, it is a
minor detail compared with detailing what methods were used and why.

A frequent theme in the boilerplate statistical methods is the
definition of statistical significance, nearly always using a p-value at
the 5\% level. This widespread use of statistical significance is
troubling giving the bright-line thinking it engenders
\citep{McShane2019} and the common misinterpretations of p-values
\citep{Goodman2008}.

Despite the extensive array of statistical tests available, many authors
are reporting the same few methods.

One reason these inadequate sections get published is that most journals
do not use statistical reviewers, despite empirical evidence showing
they improve manuscript quality \citep{Hardwicke2020}.

A related paper has criticised vague statistical methods sections
because they deprive readers and reviewers for the opportunity to
confirm that the appropriate methods were used \citep{Weissgerber2018}.
These authors checked hundreds of papers using ANOVA and found that 95\%
did not contain the information needed to determine what type of ANOVA
was performed. This lack of information could well be because the
authors used a boilerplate statistical methods section that was missing
key details.

If authors shared their code then this would provide an alternative
route for checking what statistical methods were used. This is not a
perfect solution, as we still want authors to accurately report their
methods, but it does increase transparency. However, a recent paper
found that code sharing was very low in biomedical papers, with just 2\%
of a sample of over 6,000 papers sharing code \citep{Serghiou2021}.

Many researchers are using lazy practice by copying a standard
``boilerplate'' statistical methods section, likely cut-and-pasting from
other researchers or projects. This is a strong sign of the ritualistic
practice of statistics where researchers go through the motions rather
than using conscientious practice \citep{Stark2018}. This is concerning
because using the wrong statistical methods can reduce the value of
study, or worse, invalidate the entire study. These mistakes are
avoidable and are wasting of thousands of hours of researchers' time and
the time of patients and volunteers. Poor statistical practice is a key
driver of the ongoing reproducibility crisis in science
\citep{Ioannidis2014}.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

We did not check whether papers used the correct methods, and for some
simple studies a `boilerplate' statistical methods might be adequate.

We examined papers where there was a statistics section, and we missed
papers that used statistical analysis but did not include a statistical
analysis section. Reiterate outcomes of random sample checking here.

We only examined one large journal and one trial registry and hence our
results may not be generalisable to all journals or registries,
especially those that consistently use a statistical reviewer.

We searched the full text of \emph{PLOS ONE} papers but not the
supporting information which may contain statistical methods sections
for some papers. The search terms we used to find statistical methods
appeared in the supporting information titles for xxx papers (x\%). We
did not include the supporting information because it is less structured
than the paper and could be in PDF or Word format.

\bibliographystyle{agsm}
\bibliography{references}

\end{document}